Understanding the Lomb-Scargle Periodogram

### 2016-12-26 PracticalLombScargle.tex
Start Schuster->Lomb Scargle discussion

\begin{figure}[ht]
\centering
\includegraphics[width=\textwidth]{fig01_LINEAR_data}
\caption{Observed light curve from LINEAR object ID 11375941. Uncertainties
  are indicated by the gray errorbars on each point.
  \figlabel{LINEAR_data}
}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=\textwidth]{fig02_LINEAR_PSD}
\caption{{\it Left panel:} the Lomb-Scargle periodogram computed from the data
    shown in \fig{LINEAR_data}, with an inset detailing the region near the peak.
    {\it Right panel:} the input data in \fig{LINEAR_data}, folded over the
    detected 2.58-hour period to show the coherent periodic variability.
    For more discussion of this data, see \citep{LINEAR3}.
    \figlabel{LINEAR_power}
}
\end{figure}

The Lomb-Scargle periodogram \citep{Lomb76, Scargle82}
is a well-known algorithm for detecting periodicity
in unevenly-sampled time-series, particularly within the astronomy community.
For example, consider the data shown in \fig{LINEAR_data}.
This is an irregularly-sampled timeseries showing a single object from the
LINEAR survey \citep{LINEAR1, LINEAR3}, with magnitude measured 280 times over
the course of five and a half years\footnote{
  Python code to reproduce \fig{LINEAR_data}, as well as all other figures
  in this manuscript, is available at 
  \url{http://github.com/jakevdp/PracticalLombScargle/}}.

@@

The left panel of \fig{LINEAR_power} shows the Lomb-Scargle periodogram computed for the data in \fig{LINEAR_data}.
Computing the Lomb-Scargle periodogram for the data gives us a measure of the
power as a function of period of oscillation (\fig{LINEAR_power}, left), from
which we can determine the period of oscillation of approximately 2.58 hours.
The right panel of \fig{LINEAR_power} shows a folded visualization of
the same data as \fig{LINEAR_data} -- i.e.{} plotted as a function of phase
rather than time.\footnote{
    The periodogram in \fig{LINEAR_power} was computed using
    {\tt astropy.stats.LombScargle}
    from the AstroPy project; see \url{http://astropy.org/}.
}

Often this is exactly how the Lomb-Scargle periodogram is presented: as a clean, well-defined procedure to detect the periodic component in an unevenly-sampled dataset.
In practice, however, there are a number of subtle issues that must be considered when applying a Lomb-Scargle analysis to real-world datasets, and I have found that these are rarely presented together at an introductory level.
Here are a few questions in particular that we might wish to ask about the results in \fig{LINEAR_power}:
\begin{enumerate}
  \item What is the source of the multiple peaks revealed by the Lomb-Scargle Periodogram?
  \item What is a suitable range of frequencies to search for a given dataset?
  \item What is the Nyquist frequency for this type of analysis?
  \item How should we choose the spacing of the frequency grid for our periodogram?
  \item How should we understand the uncertainty of the computed frequency?
  \item What assumptions is the Lomb-Scargle periodogram making about the signal?
\end{enumerate}
Questions like these are discussed in various textbooks and review papers, but I have not come across any concise reference that discusses such questions in depth.
This paper seeks to fill that gap, and provide a practical, semi-technical guide to the effective use of the Lomb-Scargle method of periodic analysis.
This paper does not seek a complete or rigorous treatment of the mathematics involved, but rather seeks to develop the intuition of {\it how} to think about the issues involved.

@@

   \item[The Fourier transform of sinusoid with frequency $f_0$ is a sum of delta functions at $\pm f_0$.]
     Given the integral definition of the Dirac delta function, it is straightforward to show that
     \begin{equation}
       \mathcal{F}\{e^{2\pi f_0 t}\} = \delta(f - f_0).
       \eqlabel{delta_FT}
     \end{equation}
     Euler's formula shows how such a complex exponential
     can be expressed in terms of sines and cosines:
     \begin{equation}
       e^{2\pi f_0 t} = \cos(2\pi f_0 t) + i\sin(2\pi f_0 t).
       \eqlabel{Euler_formula}
     \end{equation}
     Using the linearity of the Fourier transform, we can take these identites
     and derive the following identities:
     \begin{eqnarray}
       \mathcal{F}\{\cos(2\pi f_0 t)\} &=& \frac{1}{2}\left[\delta(f - f_0) + \delta(f + f_0)\right]\nonumber\\
       \mathcal{F}\{\sin(2\pi f_0 t)\} &=& \frac{1}{2i}\left[\delta(f - f_0) - \delta(f + f_0)\right].
     \end{eqnarray}
     In other words, a sinusoidal signal with frequency $f_0$ has a Fourier transform consisting of a weighted sum of delta functions at $\pm f_0$.

@@

\begin{figure}[ht]
\centering
\includegraphics[width=\textwidth]{fig03_Fourier_pairs}
\caption{Visualization of important Fourier pairs.\figlabel{fourier_pairs}}
\end{figure}

We have already discussed that the Fourier transform of a complex exponential is a single delta function.
This is just one of many ``Fourier Pairs'' to keep in mind as we progress to understanding the Lomb-Scargle Periodogram.
We list a few more important pairs here (see \Fig{fourier_pairs} for a visual representation of the following pairs):

\begin{description}
  \item[The Fourier transform of a sinusoid is a pair of Delta functions.]
    (See \fig{fourier_pairs}{\it a})
    \begin{equation}
      \mathcal{F}\{\cos(2\pi f_0 t)\} = \frac{1}{2}\left[\delta(f-f_0) + \delta(f+f_0)\right]
    \end{equation}
    We saw this above, but repeat it here for completeness.

   \item[The Fourier transform of a Gaussian is a Gaussian.]
    (See \fig{fourier_pairs}{\it b})
     \begin{equation}
       \mathcal{F}\{{\rm N}(t; \sigma)\} = \frac{1}{\sqrt{2\pi\sigma^2}}{\rm N}\left(f;\frac{1}{2\pi\sigma}\right)
     \end{equation}
     The Gaussian function ${\rm N}(t, \sigma)$ is given by
     \begin{equation}
       {\rm N}(t; \sigma) \equiv \frac{1}{\sqrt{2\pi\sigma^2}}e^{-t^2/(2\sigma^2)}
     \end{equation}

  \item[The Fourier transform of a tophat function is a sinc function.]
    (See \fig{fourier_pairs}{\it c})
    \begin{equation}
      \mathcal{F}\{\Pi_T(t)\} = \sinc(f T)
    \end{equation}
    The tophat function, $\Pi_T(t)$, is a normalized symmetric function that
    is uniform within a range given by $T$, and zero elsewhere:
    \begin{equation}
      \Pi_T(t)  \equiv \left\{
      \begin{array}{ll}
        1 / T, & |t| \le T / 2 \\
        0,     & |t| > T / 2
      \end{array}
      \right.
    \end{equation}
    The sinc function is given by the standard definition:
    \begin{equation}
      \sinc(x) \equiv \frac{\sin(\pi x)}{\pi x}
    \end{equation}

  \item[The Fourier transform of a Dirac comb is a Dirac comb.]
    (See \fig{fourier_pairs}{\it d})
    \begin{equation}
      \mathcal{F}\{\III_T(t)\} = \frac{1}{T}\III_{1/T}(f)
    \end{equation}
    The Dirac comb, $\III_T(t)$, is an infinite sequence of Dirac delta functions placed at even intervals of size $T$:
    \begin{equation}
      \III_T(t) \equiv \sum_{n=-\infty}^\infty \delta(t - nT)
      \eqlabel{dirac_comb}
    \end{equation}
\end{description}
The key observation in each of these Fourier pairs is the reciprocity
of scales between a function and its Fourier transform:
that is, a function with a characteristic scale $T$ will in general
have a Fourier transform with characteristic scale of $1/T$.
This feature will turn out to be very important as we push further in
understanding the Lomb-Scargle Periodogram.

@@

A final property of the Fourier transform that we will discuss here is its
ability to convert convolutions into point-wise products.
A convolution of two functions, usually denoted with a $\ast$, is defined
as follows:
\begin{equation}
  [f \ast g](t) = \int_{-\infty}^\infty f(\tau)g(t - \tau) d\tau
  \eqlabel{convolution_definition}
\end{equation}
From the definition, it is clear that a convolution amounts to ``sliding'' one
function past the other.
Such an operation is commonly used, for example, in smoothing a function,
as visualized in \Fig{convolution} for a tophat-shaped smoothing window.

@@

Given this definition of a convolution, it can be shown that the Fourier transform of a convolution is the point-wise product of the Fourier transforms:
\begin{equation}
  \mathcal{F}\{f \ast g\} = \mathcal{F}\{f\} \cdot \mathcal{F}\{g\}
  \eqlabel{convolution_theorem}
\end{equation}
In practice, this is a much more efficient means of computing the convolution
than to directly solve at each time $t$ the integral over $\tau$ that appears
in \eq{convolution_definition}.
This result in \eq{convolution_theorem} is known as the
{\it convolution theorem}, and is illustrated in \fig{convolution_theorem}.
\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{fig05_Convolution_Theorem}
  \caption{Visualization of the convolution theorem (\eq{convolution_theorem}).
    The key is that the Fourier transform of
    a convolution is the pointwise product of the two Fourier transforms.
    In the right panels, the black and gray lines represent the real and
    imaginary part of the transform, respectively.
    \figlabel{convolution_theorem}}
\end{figure}
When thinking about frequency components of time-domain measurements, this
property of the Fourier transform becomes essential, as we will see.

An important corrollary is that the Fourier transform of a product is a convolution of the two transforms:
\begin{equation}
  \mathcal{F}\{f \cdot g\} = \mathcal{F}\{f\} \ast \mathcal{F}\{g\}
  \eqlabel{convolution_theorem_inverse}
\end{equation}
We will explore the impact of this in the next section.

@@

\subsection{Effect of a Tophat Window}
First, let's consider the case of observing a continuous periodic signal over
a limited span of time, as illustrated in \Fig{tophat_window}.
Here we consider a periodic function made up of three frequency components, observed within a 10-unit time window.
The observed signal in this case can be thought of as a pointwise product of an infinite periodic signal, and a tophat window function.
Using \eq{convolution_theorem_inverse}, we see that the Fourier transform of
this function is equivalent to the convolution of the Fourier transforms of each component, which are a set of delta functions and a sinc function, respectively.
For a purely periodic signal, this convolution has the effect of broadening the delta functions representing the signal, turning each of them into a shifted sinc function.
Because of the inverse relationship between the width of the window and the width of its transform (see \Fig{fourier_pairs}), it follows that a wider observing window leads to proportionally less spread in the Fourier transform of the observed function.

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{fig06_Tophat_Window}
  \caption{Visualization of the effect on the Fourier transform of a
    tophat-shaped observing window (i.e., a continuous signal observed
    fully, but only within a finite range of time). The observed Fourier
    transform is a convolution of the true transform (here a series of Delta
    functions) and the window transform (here a narrow sinc function).
    \figlabel{tophat_window}}
\end{figure}

\subsection{The Dirac Comb and the Discrete Fourier Transform}
Another window function that commonly arises is when a continuous signal is
sampled instantaneously at regular intervals.
Such an observation is, in effect, a point-wise product between the true
underlying signal and a Dirac comb with the $T$ parameter matching the spacing
of the observations; this is illustrated in \fig{comb_window_1}.

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{fig07_comb_window_1}
  \caption{Visualization of the effect on the Fourier transform of a
    Dirac Comb observing window (i.e., a long string of evenly-spaced
    discrete observations). The observed Fourier
    transform is a convolution of the true transform (here a localized
    Gaussian) and the window transform (here another Dirac comb).
    \figlabel{comb_window_1}}
\end{figure}

@@

The example in \fig{comb_window_1} is somewhat of a best-case scenario, because
the true Fourier transform values lie entirely in a range of width $1/T$.
If we change the sampling rate so that this is no longer the case, we will
have a situation similar to that in \fig{comb_window_2}.
With a lower sampling rate, the signal transform no longer fits within the
frequency window, and the true Fourier transform {\it cannot be recovered}
from the observed transform.

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{fig08_comb_window_2}
  \caption{Repeating the visualization from \fig{comb_window_1}, but here with
    a lower sampling rate. The result is that the Fourier transform of the
    window function (lower left) has spacing narrower than the Fourier transform
    of the signal (middle left), meaning the observed Fourier transform has
    aliasing of signals, such that not all frequency information can be
    recovered.
    \figlabel{comb_window_2}}
\end{figure}

\subsubsection{The Discrete Fourier Transform}
For regularly-spaced observations defined by a Dirac comb, the delta functions
serve to collapse the Fourier integral into a sum over periodic components.
Suppose we have a true (infinitely long and continuous) signal $g(t)$, but
we observe it only at a regular grid with spacing $T$. In this case, our
observed signal is $g_{obs} = g(t) \III_T(t)$ and its Fourier transform is
\begin{equation}
  \hat{g}_{obs}(f) = \sum_{n=-\infty}^\infty g(nT) e^{-2\pi i f n T},
\end{equation}
which follows directly from \eq{FT-def} and \eq{dirac_comb}.

In the real world, however, we will not have an infinite number of observations,
but rather a finite number of samples $N$.
We can choose the coordinate system appropriately and define
$g_n \equiv g(nT)$ to write
\begin{equation}
  \hat{g}_{obs}(f) = \sum_{n=0}^N g_n e^{-2\pi i f n T}
  \eqlabel{DFT_f}
\end{equation}
From the arguments around Nyquist aliasing, we know that the only relevant
frequency range is from $0 \le f \le 1/T$, and so we can define $N$
evenly-spaced frequencies with $\Delta f = 1 / (NT)$ covering this range.
Denoting the sampled transform as
$\hat{g}_k \equiv \hat{g}_{obs}(k\Delta f)$, we can write
\begin{equation}
  \hat{g}_k = \sum_{n=0}^N g_n e^{-2\pi i k n / N}
  \eqlabel{DFT}
\end{equation}
which you might recognize as the standard form of the discrete Fourier
transform.

You may have noticed we glossed over something here: the effect of switching
from an infinite number of samples to a finite number of samples.
In moving from $\hat{g}^{(T)}$ to $\hat{g}^{(N,T)}$, we have effectively applied
to our data a rectangular window function of width $NT$.
From the discussion accompanying \fig{tophat_window}, we know what this does:
it gives us a Fourier transform convolved with a sinc function of width
$1 / (NT)$, resulting in the ``smearing'' of the Fourier transform signal
with this width.
Roughly speaking, then, any two Fourier transform values at frequencies within
$1/(NT)$ of each other will not be independent, and so we should space our
evaluations of the frequency with $\Delta f \ge 1/(NT)$.
Comparing to above, we see that this is {\it exactly the frequency spacing}
we arrived at from Nyquist-frequency arguments.

@@

With the discrete Fourier transform defined in \eqs{DFT_f}{DFT}, we can
apply the definition of the Fourier power spectrum from \eq{power-spectrum}
to compute the {\it Schuster Periodogram}, first proposed by \citet{Schuster98}:
\begin{equation}
  P_S(f) = \frac{1}{N}\left|\sum_{n=1}^N g_n e^{-2\pi i f t_n}\right|^2
  \eqlabel{schuster-periodogram}
\end{equation}
Apart from the $1/N$ scaling, this is precisely the Fourier power defined
in \eq{power-spectrum} for the expression in \eq{DFT_f}, and it follows that
for a uniformly-sampled function $g_n$, the Schuster periodogram captures
all of the relevant frequency information present in the data.

@@

Just as in the uniform case, the observed signal is a convolution of the true
signal transform and the window transform; unlike in the uniform case, the
window transform will generally {\it not} be a nice clean function.
\Fig{random_window} shows the effect on the Fourier transform of a
non-uniform observing cadence with a sampling rate identical to that
in \fig{comb_window_1}.

A few things stand-out in this figure. In particular, the Fourier transform of
the non-uniformly spaced delta functions looks like random noise, and in some
senses it is: the Fourier space representation reflects the frequency of
observations, and so non-structured spacing of observations will lead to
a non-structured Fourier representation of the window function.
This non-structured window transform, when convolved with the Fourier transform
of the true signal, results in an observed Fourier transform with the same
lack of useful structure.
Unlike the uniform result in \Fig{comb_window_1}, there is no direct aliasing
of the true signal, and no way to exactly recover any portion of the true
Fourier transform of this data.

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{fig10_random_window}
  \caption{The effect of non-uniform sampling on the observed Fourier transform.
    These samples have the same average spacing as those in \fig{comb_window_1},
    but the lack of structure in observing window translates to a lack of
    structure in its transform, causing the observed transform to be ``noisy''.
    \figlabel{random_window}}
\end{figure}

One might hope that sampling the signal more densely might limit these problems,
and it does, but only to a degree.
\Fig{random_window_2} increases the density of observations by a factor of 10,
such that there are 200 total observations over the length-10 observing window.
The observed Fourier transform in this case is much more reflective of the
underlying signal, but still contains a degree of ``noise'' rooted in the
unstructured nature of the spacing between observations.

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{fig11_random_window_2}
  \caption{The effect of non-uniform sampling on the observed Fourier transform,
    with a factor of 10 more samples than \fig{random_window}.
    Even with very dense sampling of the function, the Fourier transform
    cannot be exactly recovered due to the 
    These samples have the same average spacing as those in \fig{comb_window_1},
    but the lack of structure in observing window translates to a lack of
    structure in its transform, causing the observed transform to be a
    ``noisy'' representation of the true transform
    \figlabel{random_window_2}}
\end{figure}

@@

\Fig{nyquist_eyer99} shows an example of such a Nyquist frequency.
The data are non-uniformly sampled at times $t_i = n_i * p$, with $p=0.01$.
This results in a Nyquist frequency $f_{Ny} = 50$, and we see the expected
behavior beyond this frequency: the signal at $f > f_{Ny}$ consists of repeated
aliases of the signal at $f < f_{Ny}$.

>>>>

\begin{figure}[ht]
\centering
\includegraphics[width=\textwidth]{fig01_LINEAR_data}
\caption{Observed light curve from LINEAR object ID 11375941. Uncertainties
  are indicated by the gray errorbars on each point.
  \figlabel{LINEAR-data}
}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=\textwidth]{fig02_LINEAR_PSD}
\caption{{\it Left panel:} the Lomb-Scargle periodogram computed from the data
    shown in \fig{LINEAR-data}, with an inset detailing the region near the peak.
    {\it Right panel:} the input data in \fig{LINEAR-data}, folded over the
    detected 2.58-hour period to show the coherent periodic variability.
    For more discussion of this data, see \citep{LINEAR3}.
    \figlabel{LINEAR-power}
}
\end{figure}

The Lomb-Scargle periodogram \citep{Lomb76, Scargle82}
is a well-known algorithm for detecting periodicity
in unevenly-sampled time-series, particularly within the astronomy community.
For example, consider the data shown in \fig{LINEAR-data}.
This is an irregularly-sampled timeseries showing a single object from the
LINEAR survey \citep{LINEAR1, LINEAR3}, with magnitude measured 280 times over
the course of five and a half years\footnote{
  Python code to reproduce \fig{LINEAR-data}, as well as all other figures
  in this manuscript, is available at 
  \url{http://github.com/jakevdp/PracticalLombScargle/}}.

@@

The left panel of \fig{LINEAR-power} shows the Lomb-Scargle periodogram computed for the data in \fig{LINEAR-data}.
Computing the Lomb-Scargle periodogram for the data gives us a measure of the
power as a function of period of oscillation (\fig{LINEAR-power}, left), from
which we can determine the period of oscillation of approximately 2.58 hours.
The right panel of \fig{LINEAR-power} shows a folded visualization of
the same data as \fig{LINEAR-data} -- i.e.{} plotted as a function of phase
rather than time.\footnote{
    The periodogram in \fig{LINEAR-power} was computed using
    {\tt astropy.stats.LombScargle}
    from the AstroPy project; see \url{http://astropy.org/}.
}

Often this is exactly how the Lomb-Scargle periodogram is presented: as a clean, well-defined procedure to detect the periodic component in an unevenly-sampled dataset.
In practice, however, there are a number of subtle issues that must be considered when applying a Lomb-Scargle analysis to real-world datasets, and I have found that these are rarely presented together at an introductory level.
Here are a few questions in particular that we might wish to ask about the results in \fig{LINEAR-power}:
\begin{enumerate}
  \item What is the source of the multiple peaks revealed by the Lomb-Scargle Periodogram?
  \item What is a suitable range of frequencies to search for a given dataset?
  \item What is the Nyquist frequency for this type of analysis?
  \item How should we choose the spacing of the frequency grid for our periodogram?
  \item How should we understand the uncertainty of the computed frequency?
  \item What assumptions is the Lomb-Scargle periodogram making about the signal?
\end{enumerate}
Questions like these are discussed in various textbooks and review papers, but I have not come across any concise reference that discusses such questions in depth.
This paper seeks to fill that gap, and provide a practical, semi-technical guide to the effective use of the Lomb-Scargle method of periodic analysis.
This paper does not seek a complete or rigorous treatment of the mathematics involved, but rather seeks to develop the intuition of {\it how} to think about the issues involved.

@@

   \item[The Fourier transform of sinusoid with frequency $f_0$ is a sum of delta functions at $\pm f_0$.]
     Given the integral definition of the Dirac delta function, it is straightforward to show that
     \begin{equation}
       \mathcal{F}\{e^{2\pi f_0 t}\} = \delta(f - f_0).
       \eqlabel{delta-FT}
     \end{equation}
     Euler's formula shows how such a complex exponential
     can be expressed in terms of sines and cosines:
     \begin{equation}
       e^{2\pi f_0 t} = \cos(2\pi f_0 t) + i\sin(2\pi f_0 t).
       \eqlabel{Euler-formula}
     \end{equation}
     Using the linearity of the Fourier transform, we can take these identites
     and derive the following identities:
     \begin{eqnarray}
       \mathcal{F}\{\cos(2\pi f_0 t)\} &=& \frac{1}{2}\left[\delta(f - f_0) + \delta(f + f_0)\right]\nonumber\\
       \mathcal{F}\{\sin(2\pi f_0 t)\} &=& \frac{1}{2i}\left[\delta(f - f_0) - \delta(f + f_0)\right].
     \end{eqnarray}
     In other words, a sinusoidal signal with frequency $f_0$ has a Fourier transform consisting of a weighted sum of delta functions at $\pm f_0$.

@@

\begin{figure}[ht]
\centering
\includegraphics[width=\textwidth]{fig03_Fourier_pairs}
\caption{Visualization of important Fourier pairs.\figlabel{fourier-pairs}}
\end{figure}

We have already discussed that the Fourier transform of a complex exponential is a single delta function.
This is just one of many ``Fourier Pairs'' to keep in mind as we progress to understanding the Lomb-Scargle Periodogram.
We list a few more important pairs here (see \Fig{fourier-pairs} for a visual representation of the following pairs):

\begin{description}
  \item[The Fourier transform of a sinusoid is a pair of Delta functions.]
    (See \fig{fourier-pairs}{\it a})
    \begin{equation}
      \mathcal{F}\{\cos(2\pi f_0 t)\} = \frac{1}{2}\left[\delta(f-f_0) + \delta(f+f_0)\right]
    \end{equation}
    We saw this above, but repeat it here for completeness.

   \item[The Fourier transform of a Gaussian is a Gaussian.]
    (See \fig{fourier-pairs}{\it b})
     \begin{equation}
       \mathcal{F}\{{\rm N}(t; \sigma)\} = \frac{1}{\sqrt{2\pi\sigma^2}}{\rm N}\left(f;\frac{1}{2\pi\sigma}\right)
     \end{equation}
     The Gaussian function ${\rm N}(t, \sigma)$ is given by
     \begin{equation}
       {\rm N}(t; \sigma) \equiv \frac{1}{\sqrt{2\pi\sigma^2}}e^{-t^2/(2\sigma^2)}
     \end{equation}

  \item[The Fourier transform of a tophat function is a sinc function.]
    (See \fig{fourier-pairs}{\it c})
    \begin{equation}
      \mathcal{F}\{\Pi_T(t)\} = \sinc(f T)
    \end{equation}
    The tophat function, $\Pi_T(t)$, is a normalized symmetric function that
    is uniform within a range given by $T$, and zero elsewhere:
    \begin{equation}
      \Pi_T(t)  \equiv \left\{
      \begin{array}{ll}
        1 / T, & |t| \le T / 2 \\
        0,     & |t| > T / 2
      \end{array}
      \right.
    \end{equation}
    The sinc function is given by the standard definition:
    \begin{equation}
      \sinc(x) \equiv \frac{\sin(\pi x)}{\pi x}
    \end{equation}

  \item[The Fourier transform of a Dirac comb is a Dirac comb.]
    (See \fig{fourier-pairs}{\it d})
    \begin{equation}
      \mathcal{F}\{\III_T(t)\} = \frac{1}{T}\III_{1/T}(f)
    \end{equation}
    The Dirac comb, $\III_T(t)$, is an infinite sequence of Dirac delta functions placed at even intervals of size $T$:
    \begin{equation}
      \III_T(t) \equiv \sum_{n=-\infty}^\infty \delta(t - nT)
      \eqlabel{dirac-comb}
    \end{equation}
\end{description}
The key observation in each of these Fourier pairs is the reciprocity
of scales between a function and its Fourier transform:
that is, a function with a characteristic scale $T$ will in general
have a Fourier transform with characteristic scale of $1/T$.
This feature will turn out to be very important as we push further in
understanding the Lomb-Scargle Periodogram.

@@

A final property of the Fourier transform that we will discuss here is its
ability to convert convolutions into point-wise products.
A convolution of two functions, usually denoted with a $\ast$, is defined
as follows:
\begin{equation}
  [f \ast g](t) = \int_{-\infty}^\infty f(\tau)g(t - \tau) d\tau
  \eqlabel{convolution-definition}
\end{equation}
From the definition, it is clear that a convolution amounts to ``sliding'' one
function past the other.
Such an operation is commonly used, for example, in smoothing a function,
as visualized in \Fig{convolution} for a tophat-shaped smoothing window.

@@

Given this definition of a convolution, it can be shown that the Fourier transform of a convolution is the point-wise product of the Fourier transforms:
\begin{equation}
  \mathcal{F}\{f \ast g\} = \mathcal{F}\{f\} \cdot \mathcal{F}\{g\}
  \eqlabel{convolution-theorem}
\end{equation}
In practice, this is a much more efficient means of computing the convolution
than to directly solve at each time $t$ the integral over $\tau$ that appears
in \eq{convolution-definition}.
This result in \eq{convolution-theorem} is known as the
{\it convolution theorem}, and is illustrated in \fig{convolution-theorem}.
\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{fig05_Convolution_Theorem}
  \caption{Visualization of the convolution theorem (\eq{convolution-theorem}).
    The key is that the Fourier transform of
    a convolution is the pointwise product of the two Fourier transforms.
    In the right panels, the black and gray lines represent the real and
    imaginary part of the transform, respectively.
    \figlabel{convolution-theorem}}
\end{figure}
When thinking about frequency components of time-domain measurements, this
property of the Fourier transform becomes essential, as we will see.

An important corrollary is that the Fourier transform of a product is a convolution of the two transforms:
\begin{equation}
  \mathcal{F}\{f \cdot g\} = \mathcal{F}\{f\} \ast \mathcal{F}\{g\}
  \eqlabel{convolution-theorem-inverse}
\end{equation}
We will explore the impact of this in the next section.

@@

\subsection{Effect of a Tophat Window}
First, let's consider the case of observing a continuous periodic signal over
a limited span of time, as illustrated in \Fig{tophat-window}.
Here we consider a periodic function made up of three frequency components, observed within a 10-unit time window.
The observed signal in this case can be thought of as a pointwise product of an infinite periodic signal, and a tophat window function.
Using \eq{convolution-theorem-inverse}, we see that the Fourier transform of
this function is equivalent to the convolution of the Fourier transforms of each component, which are a set of delta functions and a sinc function, respectively.
For a purely periodic signal, this convolution has the effect of broadening the delta functions representing the signal, turning each of them into a shifted sinc function.
Because of the inverse relationship between the width of the window and the width of its transform (see \Fig{fourier-pairs}), it follows that a wider observing window leads to proportionally less spread in the Fourier transform of the observed function.

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{fig06_Tophat_Window}
  \caption{Visualization of the effect on the Fourier transform of a
    tophat-shaped observing window (i.e., a continuous signal observed
    fully, but only within a finite range of time). The observed Fourier
    transform is a convolution of the true transform (here a series of Delta
    functions) and the window transform (here a narrow sinc function).
    \figlabel{tophat-window}}
\end{figure}

\subsection{The Dirac Comb and the Discrete Fourier Transform}
Another window function that commonly arises is when a continuous signal is
sampled instantaneously at regular intervals.
Such an observation is, in effect, a point-wise product between the true
underlying signal and a Dirac comb with the $T$ parameter matching the spacing
of the observations; this is illustrated in \fig{comb-window-1}.

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{fig07_comb_window_1}
  \caption{Visualization of the effect on the Fourier transform of a
    Dirac Comb observing window (i.e., a long string of evenly-spaced
    discrete observations). The observed Fourier
    transform is a convolution of the true transform (here a localized
    Gaussian) and the window transform (here another Dirac comb).
    \figlabel{comb-window-1}}
\end{figure}

@@

The example in \fig{comb-window-1} is somewhat of a best-case scenario, because
the true Fourier transform values lie entirely in a range of width $1/T$.
If we change the sampling rate so that this is no longer the case, we will
have a situation similar to that in \fig{comb-window-2}.
With a lower sampling rate, the signal transform no longer fits within the
frequency window, and the true Fourier transform {\it cannot be recovered}
from the observed transform.

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{fig08_comb_window_2}
  \caption{Repeating the visualization from \fig{comb-window-1}, but here with
    a lower sampling rate. The result is that the Fourier transform of the
    window function (lower left) has spacing narrower than the Fourier transform
    of the signal (middle left), meaning the observed Fourier transform has
    aliasing of signals, such that not all frequency information can be
    recovered.
    \figlabel{comb-window-2}}
\end{figure}

\subsubsection{The Discrete Fourier Transform}
For regularly-spaced observations defined by a Dirac comb, the delta functions
serve to collapse the Fourier integral into a sum over periodic components.
Suppose we have a true (infinitely long and continuous) signal $g(t)$, but
we observe it only at a regular grid with spacing $T$. In this case, our
observed signal is $g_{obs} = g(t) \III_T(t)$ and its Fourier transform is
\begin{equation}
  \hat{g}_{obs}(f) = \sum_{n=-\infty}^\infty g(nT) e^{-2\pi i f n T},
\end{equation}
which follows directly from \eq{FT-def} and \eq{dirac-comb}.

In the real world, however, we will not have an infinite number of observations,
but rather a finite number of samples $N$.
We can choose the coordinate system appropriately and define
$g_n \equiv g(nT)$ to write
\begin{equation}
  \hat{g}_{obs}(f) = \sum_{n=0}^N g_n e^{-2\pi i f n T}
  \eqlabel{DFT-f}
\end{equation}
From the arguments around Nyquist aliasing, we know that the only relevant
frequency range is from $0 \le f \le 1/T$, and so we can define $N$
evenly-spaced frequencies with $\Delta f = 1 / (NT)$ covering this range.
Denoting the sampled transform as
$\hat{g}_k \equiv \hat{g}_{obs}(k\Delta f)$, we can write
\begin{equation}
  \hat{g}_k = \sum_{n=0}^N g_n e^{-2\pi i k n / N}
  \eqlabel{DFT}
\end{equation}
which you might recognize as the standard form of the discrete Fourier
transform.

You may have noticed we glossed over something here: the effect of switching
from an infinite number of samples to a finite number of samples.
In moving from $\hat{g}^{(T)}$ to $\hat{g}^{(N,T)}$, we have effectively applied
to our data a rectangular window function of width $NT$.
From the discussion accompanying \fig{tophat-window}, we know what this does:
it gives us a Fourier transform convolved with a sinc function of width
$1 / (NT)$, resulting in the ``smearing'' of the Fourier transform signal
with this width.
Roughly speaking, then, any two Fourier transform values at frequencies within
$1/(NT)$ of each other will not be independent, and so we should space our
evaluations of the frequency with $\Delta f \ge 1/(NT)$.
Comparing to above, we see that this is {\it exactly the frequency spacing}
we arrived at from Nyquist-frequency arguments.

@@

With the discrete Fourier transform defined in \eqs{DFT-f}{DFT}, we can
apply the definition of the Fourier power spectrum from \eq{power-spectrum}
to compute the {\it Schuster Periodogram}, first proposed by \citet{Schuster98}:
\begin{equation}
  P_S(f) = \frac{1}{N}\left|\sum_{n=1}^N g_n e^{-2\pi i f t_n}\right|^2
  \eqlabel{schuster-periodogram}
\end{equation}
Apart from the $1/N$ scaling, this is precisely the Fourier power defined
in \eq{power-spectrum} for the expression in \eq{DFT-f}, and it follows that
for a uniformly-sampled function $g_n$, the Schuster periodogram captures
all of the relevant frequency information present in the data.

@@

Just as in the uniform case, the observed signal is a convolution of the true
signal transform and the window transform; unlike in the uniform case, the
window transform will generally {\it not} be a nice clean function.
\Fig{random-window} shows the effect on the Fourier transform of a
non-uniform observing cadence with a sampling rate identical to that
in \fig{comb-window-1}.

A few things stand-out in this figure. In particular, the Fourier transform of
the non-uniformly spaced delta functions looks like random noise, and in some
senses it is: the Fourier space representation reflects the frequency of
observations, and so non-structured spacing of observations will lead to
a non-structured Fourier representation of the window function.
This non-structured window transform, when convolved with the Fourier transform
of the true signal, results in an observed Fourier transform with the same
lack of useful structure.
Unlike the uniform result in \Fig{comb-window-1}, there is no direct aliasing
of the true signal, and no way to exactly recover any portion of the true
Fourier transform of this data.

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{fig10_random_window}
  \caption{The effect of non-uniform sampling on the observed Fourier transform.
    These samples have the same average spacing as those in \fig{comb-window-1},
    but the lack of structure in observing window translates to a lack of
    structure in its transform, causing the observed transform to be ``noisy''.
    \figlabel{random-window}}
\end{figure}

One might hope that sampling the signal more densely might limit these problems,
and it does, but only to a degree.
\Fig{random-window-2} increases the density of observations by a factor of 10,
such that there are 200 total observations over the length-10 observing window.
The observed Fourier transform in this case is much more reflective of the
underlying signal, but still contains a degree of ``noise'' rooted in the
unstructured nature of the spacing between observations.

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{fig11_random_window_2}
  \caption{The effect of non-uniform sampling on the observed Fourier transform,
    with a factor of 10 more samples than \fig{random-window}.
    Even with very dense sampling of the function, the Fourier transform
    cannot be exactly recovered due to the 
    These samples have the same average spacing as those in \fig{comb-window-1},
    but the lack of structure in observing window translates to a lack of
    structure in its transform, causing the observed transform to be a
    ``noisy'' representation of the true transform
    \figlabel{random-window-2}}
\end{figure}

@@

\Fig{nyquist-eyer99} shows an example of such a Nyquist frequency.
The data are non-uniformly sampled at times $t_i = n_i * p$, with $p=0.01$.
This results in a Nyquist frequency $f_{Ny} = 50$, and we see the expected
behavior beyond this frequency: the signal at $f > f_{Ny}$ consists of repeated
aliases of the signal at $f < f_{Ny}$.
